---
title: "AustinNebel"
author: "Austin Nebel"
date: "4/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(readr)
library(readxl)
library(dplyr)
library(devtools)
```



```{r}
data <- read.csv("healthcare-dataset-stroke-data.csv")
summary(data)
```

## How does working environment and residence type affect stroke chances?

```{r}
data %>% group_by(work_type, Residence_type) %>%
  mutate(stroke_avg = sum(stroke)/n()) %>%
  ggplot(aes(x = `work_type`, y=`stroke_avg`, fill = `Residence_type`)) + geom_bar(stat="identity", position = "dodge")
```

The above bar graph shows that working environment plays a large role in the chance of a stroke, and even residence type can be a small factor. Those living in urban households consistently had slightly more strokes that those in rural households. What is more compelling is that the working environment significantly changed the amount of strokes; those that are self-employed had nearly a 1 in 10 chance of having a stroke. On the contrary, those that stayed home to raise children had less than 0.5% chance of developing a stroke in their lifetime.
