---
title: "Final Project"
author: "Ryan Herren"
date: "4/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(readr)
library(readxl)
library(dplyr)
library(devtools)
library(qwraps2)
library(expss)
```

```{r}
data <- read.csv("/Users/ryanherren/Desktop/DS202/FinalProject/healthcare-dataset-stroke-data.csv")
head(data)
```

## This chart explores the trend of strokes vs peoples ages.
## Author: Ryan Herren
```{r}
strokes <- data[which(data$stroke == 1 & data$bmi != 'N/A'), ]
ggplot(strokes, aes(x = age)) + geom_bar(color = 'red') + xlab("Age") + ylab("Number of Strokes") + ggtitle("Number of Strokes by Age")
```

## This chart explores the bmi and age of all people who reported having a stroke.
## Author: Ryan Herren
```{r}
ggplot(strokes, aes(x=age, y = as.numeric(bmi))) + geom_point() + labs(x = "Age", y = "BMI")
```

## This chart explores the bmi and age of all people in the data set, whether or not they've had a stroke before. 
## Author: Ryan Herren
```{r}
data1 <- na.omit(data)
data1 <- data1[which(data1$bmi != 'N/A'), ]
vstroke <- data1[which(data1$stroke == 1), ]
vnostroke <- data1[which(data1$stroke == 0) ,]
p1 <- ggplot() + geom_point(vnostroke, mapping = aes(x = age, y = as.numeric(bmi))) + geom_point(vstroke, mapping = aes(x = age, y = as.numeric(bmi), colour = 'red'))
p1 + labs(x = "Age", y = "BMI") + ggtitle("Age vs Body Mass Index \nAll individuals who have had a stroke are marked by a red dot") + theme(legend.position = "none")
```

## This chart explores the relationship between Average Glucose Level, BMI, and stroke occurences.  
## Author: Ryan Herren
```{r}
AGLBMI <- na.omit(data)
AGLBMI <- AGLBMI[which(AGLBMI$bmi != 'N/A' & AGLBMI$avg_glucose_level != 'N/A'), ]
AGLBMIS <- AGLBMI[which(data1$stroke == 1), ]
AGLBMINS <- AGLBMI[which(data1$stroke == 0), ]
p2 <- ggplot() + geom_point(AGLBMINS, mapping = aes(x = avg_glucose_level, y = as.numeric(bmi))) + geom_point(AGLBMIS, mapping = aes(x = avg_glucose_level, y = as.numeric(bmi), colour = 'red'))
p2 + labs(x = "Average Glucose Level", y = "BMI") + ggtitle("Average Glucose Level vs Body Mass Index \nAll individuals who have had a stroke are marked by a red dot") + theme(legend.position = "none")
```
```{r, echo=FALSE}
no_strokes <- data[which(data$stroke == 0 & data$bmi != "N/A"), ]

sma <- mean(strokes[["age"]])
nma <- mean(no_strokes[["age"]])

smh <- mean(strokes[["hypertension"]])
nmh <- mean(no_strokes[["hypertension"]])

smhd <- mean(strokes[["heart_disease"]])
nmhd <- mean(no_strokes[["heart_disease"]])

smagl <- mean(strokes[["avg_glucose_level"]])
nmagl <- mean(no_strokes[["avg_glucose_level"]])

smbmi <- mean(as.numeric(strokes[["bmi"]]))
nmbmi <- mean(as.numeric(no_strokes[["bmi"]]))
```

Below is simple mean values from dividing the dataset into two separate ones - one with stroke patients, one without. 

<table>
<tr>
<td>Statistic     <br></td>
<td>Stroke Patient     <br></td>
<td>Non-Stroke Patient     <br></td>
</tr>
<tr>
<td>Age</td>
<td>67.71292</td>
<td>41.97154</td>
</tr>
<tr>
<td>Probability of Hypertension</td>
<td>0.2870813</td>
<td>0.08319149</td>
</tr>
<tr>
<td>Probability of Heart Disease</td>
<td>0.1913876</td>
<td>0.04319149</td>
</tr>
<tr>
<td>Average Glucose Level</td>
<td>134.5714</td>
<td>104.0037</td>
</tr>
<tr>
<td>BMI</td>
<td>30.47129</td>
<td>28.82306</td>
</tr>
</table>

  
  This table shows a lot of valuable information. The average stroke patient is 26 years older than the average non-stroke patient, 20% more likely to have hypertension, 15% more likely to have heart disease, has an average glucose level 30 points higher than non-stroke patients, and is 1.6 higher on the BMI scale.
  This can lead to drawing many significant solutions. Age is obviously a huge factor in your likelihood to have a stroke. Followed by hypertension and heart disease, which both put you at a higher risk of a stroke. The average glucose level is definitely a factor as well, shown by the 30 point increase on average for a stroke patient. The least significant result here is Body Mass Index. The average body mass index difference between stroke and non-stroke patients is just 1.65. This is a small number on the BMI scale, and both of these values actually lie on the upper end of the overweight category of the BMI scale.
